---
title: "National Football League"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{National Football League}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

```{r setup}
library(nflexplore)
library(dplyr)
library(ggplot2)
library(tidyr)
library(scales)
```

# Introduction to nflexplore

This package is created to explore the world of National Football League (NFL). NFL is the premier professional American football league in the United States, consisting of 32 teams divided into two conferences and four division each. It is known for its physical gameplay, complex rules, and culminates in the Super Bowl, one of the world's most-watched sporting events. The package offers users access to fundamental NFL facts and engaging datasets for game attendance and results. It also includes convenient functions to launch an interactive Shiny App for NFL statistics and plot winning trend graphs for individual NFL teams with ease. There is also additional function that will help you calculate the mean for each level of the categorical variable.

## Datasets and Use Case

This document introduces you to nflexplore's usage. For instance, there is two sample datasets that can be called, they are:

`attendance` : Dataset that shows the total attendance of NFL game per week and per team, spanning multiple years from 2000 until 2019.

`games` : Dataset that contains information about weekly games, providing details about the winning team and game performances.

By utilising these two datasets, we will be able to perform a simple data analysis that will add a new knowledge about NFL in general. Let's create some example:

### 1. Plotting a yearly attendance (using `attendance` dataset)

```{r}
# Create a plot by summarising attendance number into yearly basis
attendance %>%
    filter(week == 1) %>%
    group_by(year) %>%
    summarise(total = sum(total),
              home = sum(home)) %>%
    ggplot(aes(x = year, y = total, group = 1)) +
      geom_line(size = 1) +
      geom_point(size = 3) +
      theme_classic() +
      scale_y_continuous(labels = scales::comma)
      
```

### 2. Top 5 team with most win (using `games` dataset)

```{r plot2, message=FALSE, warning=FALSE}
#Creating home/away win table
home_away <- games %>%
                 group_by(year, home_team) %>%
                 summarise(home = sum(home_win),
                           away = sum(away_win)) %>%
                 pivot_longer(c("home","away"),
                              names_to = "home_aways",
                              values_to = "win_number")

#Summarising home/away win table
home_away_summ <- home_away %>%
                    group_by(home_team) %>%
                    summarise(total_win = sum(win_number))
                    
#Plotting for Home & Away Win
home_away_summ %>%
    slice_max(total_win, n = 5) %>%
    ggplot(aes(x = reorder(home_team, total_win), y = total_win)) +
      geom_bar(position = "dodge", stat = "identity") +
      coord_flip() +
      xlab("Team") +
      ylab("Number of Win") +
      theme_classic()
      
```

## Functions and Use Case

Apart from dataset, this package also include some functions. There are three functions that serves to increase your knowledge on NFL, which are

`run_app`: This function will open a Shiny App that has been created by the author, which provides preliminary knowledge about NFL.

`win_trend`: This function will return a home/away winning trend of a certain NFL team.

`calculate_means`: This function will calculate the mean for each level of the categorical variable and returns a named numeric vector with the means.

Let's explore the ability of each function:

### 1. `run_app` function

Whenever you call this function, it will open up a Shiny App which provides preliminary knowledge about NFL. This is an interactive web application, you can play around with the plot. For example, you can change the filter provided in the application, and the plot will change accordingly. In addition to that, there is a short video explaining how NFL games is played, including the game structure, how the teams score the points, and rules of the game.

```{r run-app, eval=FALSE}
run_app()
```

### 2. `win_trend` function

To run this function, you need to prepare one NFL team that you want to see their winning trend. You can refer to the NFL official website for the complete list of the team [here](https://www.nfl.com/teams/). Once you have identified what team that you want to explore, you can input it into the function. Here is the example:

```{r}
win_trend("Dallas Cowboys")
```

### 3. `calculate_means` function

This function takes a numeric vector and a categorical vector as input. It calculates the mean for each level of the categorical variable and returns a named numeric vector with the means. Here is the example how to use this:

```{r}
numeric_data <- c(10, 15, 20, 25, 30, 35)
category_data <- c("A", "B", "A", "B", "A", "B")

calculate_means(numeric_data, category_data)
```
